---
title: "CART"
output:
  html_document: default
  pdf_document: default
---


```{r include=FALSE}
#Replace the below command to point to your working directory
setwd ("/Users/kumar/Dropbox/_Projects/GreatLearning/GL-ML-SlidesNotes/Week2/code")
```


```{r}
CD.sample = read.csv("trainSample.csv", sep = ",", header = TRUE)
CD.holdout = read.table("testSample.csv", sep = ",", header = TRUE)
CD.sample = CD.sample[,-1]
nrow(CD.sample)
nrow(CD.holdout)
head(CD.sample)
table(CD.sample$Target)
sum(CD.sample$Target)/nrow(CD.sample)
```



```{r}

library(rpart)
library(rpart.plot)


tree1 <- rpart(formula = Target ~ ., 
            data = CD.sample, method = "class")
rpart.plot(tree1)
```


```{r}


#?rpart.control

r.ctrl = rpart.control(minsplit=100, minbucket = 10, cp = 0, xval = 10)
tree2 <- rpart(formula = Target ~ ., 
            data = CD.sample, method = "class", 
            control = r.ctrl)
rpart.plot(tree2)

```

```{r}
printcp(tree2)
plotcp(tree2)
```

```{r}
tree3 = prune(tree2, cp= 0.0021 ,"CP")
printcp(tree3)
rpart.plot(tree3)

```



## Let's use rattle to see various model evaluation measures
##rattle()

View(CD.sample)
## Scoring syntax
?predict
CD.sample$predict.class <- predict(ptree, CD.sample, type="class")
CD.sample$predict.score <- predict(ptree, CD.sample, type="prob")


View(CD.sample)
head(CD.sample)


## deciling code
decile <- function(x){
  deciles <- vector(length=10)
  for (i in seq(0.1,1,.1)){
    deciles[i*10] <- quantile(x, i, na.rm=T)
  }
  return (
  ifelse(x<deciles[1], 1,
  ifelse(x<deciles[2], 2,
  ifelse(x<deciles[3], 3,
  ifelse(x<deciles[4], 4,
  ifelse(x<deciles[5], 5,
  ifelse(x<deciles[6], 6,
  ifelse(x<deciles[7], 7,
  ifelse(x<deciles[8], 8,
  ifelse(x<deciles[9], 9, 10
  ))))))))))
}

class(CD.sample$predict.score)
## deciling
CD.sample$deciles <- decile(CD.sample$predict.score[,2])
View(CD.sample)

## Ranking code
##install.packages("data.table")
##install.packages("scales")
library(data.table)
library(scales)
tmp_DT = data.table(CD.sample)
rank <- tmp_DT[, list(
  cnt = length(Target), 
  cnt_resp = sum(Target), 
  cnt_non_resp = sum(Target == 0)) , 
  by=deciles][order(-deciles)]
rank$rrate <- round(rank$cnt_resp / rank$cnt,4);
rank$cum_resp <- cumsum(rank$cnt_resp)
rank$cum_non_resp <- cumsum(rank$cnt_non_resp)
rank$cum_rel_resp <- round(rank$cum_resp / sum(rank$cnt_resp),4);
rank$cum_rel_non_resp <- round(rank$cum_non_resp / sum(rank$cnt_non_resp),4);
rank$ks <- abs(rank$cum_rel_resp - rank$cum_rel_non_resp) * 100;
rank$rrate <- percent(rank$rrate)
rank$cum_rel_resp <- percent(rank$cum_rel_resp)
rank$cum_rel_non_resp <- percent(rank$cum_rel_non_resp)

View(rank)



##install.packages("ROCR")
##install.packages("ineq")
library(ROCR)
library(ineq)
pred <- prediction(CD.sample$predict.score[,2], CD.sample$Target)
perf <- performance(pred, "tpr", "fpr")
plot(perf)
KS <- max(attr(perf, 'y.values')[[1]]-attr(perf, 'x.values')[[1]])
auc <- performance(pred,"auc"); 
auc <- as.numeric(auc@y.values)

gini = ineq(CD.sample$predict.score[,2], type="Gini")

with(CD.sample, table(Target, predict.class))
auc
KS
gini

View(rank)

## Syntax to get the node path
tree.path <- path.rpart(ptree, node = c(2, 12))
nrow(CD.holdout)

## Scoring Holdout sample
CD.holdout$predict.class <- predict(ptree, CD.holdout, type="class")
CD.holdout$predict.score <- predict(ptree, CD.holdout, type="prob")


CD.holdout$deciles <- decile(CD.holdout$predict.score[,2])
View(CD.holdout)

## Ranking code
tmp_DT = data.table(CD.holdout)
h_rank <- tmp_DT[, list(
  cnt = length(Target), 
  cnt_resp = sum(Target), 
  cnt_non_resp = sum(Target == 0)) , 
  by=deciles][order(-deciles)]
h_rank$rrate <- round(h_rank$cnt_resp / h_rank$cnt,4);
h_rank$cum_resp <- cumsum(h_rank$cnt_resp)
h_rank$cum_non_resp <- cumsum(h_rank$cnt_non_resp)
h_rank$cum_rel_resp <- round(h_rank$cum_resp / sum(h_rank$cnt_resp),4);
h_rank$cum_rel_non_resp <- round(h_rank$cum_non_resp / sum(h_rank$cnt_non_resp),4);
h_rank$ks <- abs(h_rank$cum_rel_resp - h_rank$cum_rel_non_resp)*100;
h_rank$rrate <- percent(h_rank$rrate)
h_rank$cum_rel_resp <- percent(h_rank$cum_rel_resp)
h_rank$cum_rel_non_resp <- percent(h_rank$cum_rel_non_resp)

View(h_rank)

pred <- prediction(CD.holdout$predict.score[,2], CD.holdout$Target)
perf <- performance(pred, "tpr", "fpr")
KS <- max(attr(perf, 'y.values')[[1]]-attr(perf, 'x.values')[[1]])
auc <- performance(pred,"auc"); 
auc <- as.numeric(auc@y.values)

gini = ineq(CD.holdout$predict.score[,2], type="Gini")

with(CD.holdout, table(Target, predict.class))
auc
KS
gini


