---
title: "Model Perfomance Measures"
output:
  html_document: default
  pdf_document: default
---

```{r include=FALSE}
#Replace the below command to point to your working directory
setwd ("/Users/kumar/Dropbox/_Projects/GreatLearning/GL-ML-SlidesNotes/Week2/code")
```

Lets first load a training data set:
```{r}
trainDS = read.csv("trainSample.csv", sep = ",", header = TRUE)
#trainDS$Target=as.factor(trainDS$Target)
trainDS=trainDS[,-1]
head(trainDS)
nrow(trainDS)
sum(trainDS$Target==1)/nrow(trainDS)
```

Next, fit a simple Decision Tree model and also to use the decision tree to predict the class for each row and/or score the probabilities.

```{r}
library(rpart)
library(rpart.plot)

tree <- rpart(formula = Target ~ ., 
            data = trainDS, method = "class", minsplit=100, minbucket = 10, cp = 0.0021, xval = 10)
printcp(tree)
rpart.plot(tree)

trainDS$prediction = predict(tree, trainDS, type="class")
trainDS$prob1 = predict(tree, trainDS, type="prob")[,"1"]
head(trainDS)
```



In the rest of this document, we will sequentially go through a bunch of performance measures. Lets begin with the confusion matrix on the actual vs the predicted Target from the decision tree. The code below also calculates the error rate

```{r}
tbl=table(trainDS$Target, trainDS$prediction)
print(tbl)

print((tbl[1,2]+tbl[2,1])/14000)
```
Next lets calculate the decile thresholds and use those thresholds to compute various columns in a rank order table:
```{r}
probs=seq(0,1,length=11)
qs=quantile(trainDS$prob1, probs)
trainDS$deciles=cut(trainDS$prob1, unique(qs),include.lowest = TRUE,right=FALSE)
table(trainDS$deciles)

library(data.table)
trainDT = data.table(trainDS)
rankTbl = trainDT[, list(
  cnt = length(Target), 
  cnt_tar1 = sum(Target), 
  cnt_tar0 = sum(Target == 0)
  ), 
  by=deciles][order(-deciles)]

rankTbl$rrate = round(rankTbl$cnt_tar1 / rankTbl$cnt,4)*100;
rankTbl$cum_resp = cumsum(rankTbl$cnt_tar1)
rankTbl$cum_non_resp = cumsum(rankTbl$cnt_tar0)
rankTbl$cum_rel_resp = round(rankTbl$cum_resp / sum(rankTbl$cnt_tar1),4)*100;
rankTbl$cum_rel_non_resp = round(rankTbl$cum_non_resp / sum(rankTbl$cnt_tar0),4)*100;
rankTbl$ks = abs(rankTbl$cum_rel_resp - rankTbl$cum_rel_non_resp);

print(rankTbl)

```


We will next use the ROCR and ineq packages to compute AUC, KS and gini
```{r}
library(ROCR)
library(ineq)
predObj = prediction(trainDS$prob1, trainDS$Target)
perf = performance(predObj, "tpr", "fpr")
plot(perf)
KS = max(perf@y.values[[1]]-perf@x.values[[1]])
auc = performance(predObj,"auc"); 
auc = as.numeric(auc@y.values)

gini = ineq(trainDS$prob1, type="Gini")

```

Finally, we use the Concordance function in the InformationValue package to find the concordance and discordcance ratios:
```{r}
#install.packages('InformationValue)
library(InformationValue)
Concordance(actuals=trainDS$Target, predictedScores=trainDS$prob1)
```




